Exploring mathematical objects from custom-tailored mathematical universes

### 2019-01-09 paper-filmat.tex
More on variants of the effective topos

@@

Firstly, in place of ordinary Turing machines, one can employ the
\emph{infinite-time Turing machines} pioneered by Hamkins and
Lewis~\cite{hamkins-lewis:ittm}. These machines model \emph{hypercomputation}
in that they can run for ``longer than infinity''; more precisely, the
computational steps are indexed by the ordinal numbers instead of the natural
numbers.

@@

A second ......

>>>>

\renewcommand{\paragraph}[1]{\noindent\textbf{#1.}}

@@

\paragraph{Hypercomputation}
Firstly, in place of ordinary Turing machines, one can employ the
\emph{infinite-time Turing machines} pioneered by Hamkins and
Lewis~\cite{hamkins-lewis:ittm}. These machines model \emph{hypercomputation}
in that they can run for ``longer than infinity''; more precisely, the
computational steps are indexed by the ordinal numbers instead of the natural
numbers. For instance, an infinite-time Turing machine can trivially decide the
twin prime conjecture, by simply walking along the natural number line and
recording any twin primes it finds. Then, on day~$\omega$, it can observe
whether it has found finitely many twins or not.

@@

\paragraph{Machines of the real physical world} A second variant of the
\effective topos is obtained by using machines of the real physical world
instead of abstract Turing machines in its construction. In doing so, we of
course leave the realm of mathematics, as real-world machines are not objects
of mathematical studies, but still it is interesting to see which commitments
about the natural of the physical world imply which internal statements of the
resulting topos.

For instance, Andrej Bauer proved that inside this topos any function~$\RR \to
\RR$ is continuous if, in the physical world, only finitely many computational
steps can be carried out in finite time and if it is possible to form
tamper-free private communication channels~\cite{XXX}.  % XXX: look up precise formulation
